\propertitle

### 2019-07-08 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-07-08T16:40:48+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The maximum-entropy method has often been applied in the neurosciences to
questions of sufficient statistics. For example: does knowledge about
third-order correlations change our state of uncertainty about network
activity in an appreciable way, compared to knowledge about second-order
correlations only? The method is used as follows to answer this question.
We imagine to be ignorant of the measured frequency distribution for the
network activity, and to know only up to second-order measured statistics,
or up to third-order. Two maximum-entropy distributions are then built,
constrained under the two sets of moments under comparison. Finally, the
resulting distributions are compared with each other and with the measured
frequency distribution. This recipe generalizes to any subsets of measured
statistics, of course.

The method introduced in the present work reveals an issue in this recipe;
the issue doesn't come from the method itself but from a simple, general
fact of probability theory. If a probability distribution has some
sufficient statistics \citep[\sect~4.5]{bernardoetal1994}{fortinietal2000},
then its marginals \emph{cannot} have the same sufficient statistics, and
vice versa, except for trivial cases such as uniform distributions. This in
general also holds between different marginals: if one has some sufficient
statistics, another won't have it. This impossibility is known in
statistical mechanics: marginals of Gibbs states aren't Gibbs states
\citep{maesetal1999}. Mathematically it corresponds to the general
impossibility of solving a system of $n$ equations in $m$ unknowns when
$n>m$ \citep[\sect~3.1]{portamanaetal2015}[\sect~3]{portamanaetal2018b}.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-07-08T23:05:19+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

In the neurosciences, the maximum-entropy method has often been applied to
questions of sufficient statistics. In this section we discuss how our
proposed application bears on this kind of questions. For concreteness
we use an example with specific statistics: first and second moments
(equivalent to mean and correlations), versus first-to-fourth moments. But
our discussion holds for any other sets of statistics.

The question is as follows. We want to assess our degree of belief about
the frequencies of the activities of the sample, or about the sample
activity in a new time bin. For this assessment we use any measured data we
have. But it sometimes happens that dropping part of the data -- for
example, the measured third- and higher-order moments -- leaves our
assessment almost unchanged: the dropped data are \emph{informationally
  irrelevant}, or \emph{almost} so. The data that remain -- for example,
first and second moments -- are \emph{informationally sufficient}. This
informational quasi-sufficiency may be connected to biological or physical
mechanisms; that's why it's interesting.

We can approximately check which data are irrelevant by simply dropping
them from the conditional of our degree of belief, and see if the latter
changes appreciably. When our degree of belief is built with the
maximum-entropy method, we simply drop some of the data used as constraints
and see how the resulting distribution changes. In the neurosciences this
has been done with maximum-entropy distributions constructed \emph{for the
  sample}.

% For example: does knowledge about
% third-order correlations change our state of uncertainty about network
% activity in an appreciable way, compared to knowledge about second-order
% correlations only? The method is used as follows to answer this question.
% We imagine to be ignorant of the measured frequency distribution for the
% network activity, and to know only up to second-order measured statistics,
% or up to third-order. Two maximum-entropy distributions are then built,
% constrained under the two sets of moments under comparison. Finally, the
% resulting distributions are compared with each other and with the measured
% frequency distribution. This recipe generalizes to any subsets of measured
% statistics, of course.

The method introduced in the present work reveals an issue with this
application to the sample. The issue doesn't come from the method itself
but from a simple, general fact of probability theory. If a probability
distribution has some sufficient statistics
\citep[\sect~4.5]{bernardoetal1994}{fortinietal2000}, then its marginals
\emph{cannot} have the same sufficient statistics, and vice versa, except
for trivial cases such as uniform distributions. This in general also holds
between different marginals: if a marginal has some sufficient statistics,
another won't have it. This impossibility is known in statistical
mechanics: marginals of Gibbs states aren't Gibbs states
\citep{maesetal1999}. Mathematically it corresponds to the general
impossibility of solving a system of $n$ equations in $m$ unknowns when
$n>m$ \citep[\sect~3.1]{portamanaetal2015}[\sect~3]{portamanaetal2018b}.
